### 2024-02-23 code/data_cleaning_code.qmd
Worked on Project Part 2
Created the project part 2 document, and two code files for data cleaning and exploratory analyses. Figures and other outputs are also included

---
title: "Data cleaning"
subtitle: "Project: Part 2"
author: Erick E. Mollinedo
date: '`r format(Sys.Date(), "%B %d, %Y")`'
format: html
editor: visual
---

Packages needed for data cleaning and data exploration

```{r}
library(here)
library(readxl)
library(tidyverse)
```

## HAPIN Guatemala data cleaning

First I will load the HAPIN Guatemala dataset that has some of the variables of interest and assign it to the `hapin` dataframe, and I will also load a file that has all the filter IDs that I need to find in the `hapin` dataframe, I'll assign it to the `xrflist` dataframe.

```{r}
hapin <- read_csv(here("data", "raw-data", "HAPIN_EX17_Guatemala_20230728.csv"))
xrflist <- read_csv(here("data", "raw-data", "sa_filters_emollinedo.csv"))
```

Now I will only select the variables I am interested to work with.

```{r}
hapin <- hapin %>% select(trt_blinded, timepoint, h41_m_ecm_fid, h41_b_ecm_fid, h41b_filter1, h41b_filter2, h42_cook, h42_smoke, h42_coil, h42_trash, h42_kero, h42_incence, h42_generator, h42_smoked, h42_crop, h42_smoky_kitc, ECM_grav_neph_conc_M, ECM_bc_conc_M, CO_avg_ppm_M)
```

And now I will select only the observations from `hapin` that correspond to the filter IDs from the `xrflist` dataframe. These filter IDs could be located in either of the following variables: `h41_m_ecm_fid`, `h41_b_ecm_fid`, `h41b_filter1`, `h41b_filter2`, so I searched in these columns. But first I made sure all the filter IDs were in the same format as the IDs from the `xrflist`.

```{r}
#Make sure all filter IDs are upper case, using the `mutate()` function with `toupper()`
hapin <- hapin %>% 
  mutate(h41_m_ecm_fid = toupper(h41_m_ecm_fid),
         h41b_filter1 = toupper(h41b_filter1),
         h41_b_ecm_fid = toupper(h41_b_ecm_fid),
         h41b_filter2 = toupper(h41b_filter2))

#Select all filters that match with the list of filters from XRF, using the `filter()` function.
hapin <- hapin %>%
  filter(h41_m_ecm_fid %in% xrflist$filter_id |
           h41b_filter1 %in% xrflist$filter_id |
           h41_b_ecm_fid %in% xrflist$filter_id |
           h41b_filter2 %in% xrflist$filter_id)
```

Here I will create a new variable called `fueltype`, this variable will be coded as "LPG" or "Biomass" depending on the type of visit and the treatment or intervention `trt_blinded`.

```{r}
#Concatenate `trt_blinded` and `timepoint` into a new temporary variable
hapin$concatenated <- paste(hapin$trt_blinded, hapin$timepoint)

#Create a new variable `fueltype` that categorizes all the observations by type of fuel ('R' means the visit belongs to the 'Control' treatment and 'Q' for the 'Intervention' treatment)
hapin <- mutate(hapin, fueltype = case_when(
  endsWith(concatenated, "BL") ~ "Biomass", #BL corresponds to the baseline, where all the participants had a biomass type fuel stove
  endsWith(concatenated, "R P1") ~ "Biomass",
  endsWith(concatenated, "R P2") ~ "Biomass",
  endsWith(concatenated, "Q P1") ~ "LPG",
  endsWith(concatenated, "Q P2") ~ "LPG",
  endsWith(concatenated, "R BLP1") ~ "Biomass",
  endsWith(concatenated, "R P1P2") ~ "Biomass",
  endsWith(concatenated, "R P2B1") ~ "Biomass",
  endsWith(concatenated, "Q BLP1") ~ "LPG",
  endsWith(concatenated, "Q P1P2") ~ "LPG",
  endsWith(concatenated, "Q P2B1") ~ "LPG"))
```

And now I will rename some of the variables and delete unnecessary ones. I will also recode the values from the newly named `arm` variable.

```{r}
hapin <- hapin %>% rename(filter_id = "h41_m_ecm_fid", dup1_filterid = "h41b_filter1", dup2_filterid = "h41b_filter2",
                         blank_filterid = "h41_b_ecm_fid", arm = "trt_blinded", pm25 = "ECM_grav_neph_conc_M",
                         bc = "ECM_bc_conc_M", co = "CO_avg_ppm_M", stove = "h42_cook", smoke = "h42_smoke",
                         coil = "h42_coil", trash = "h42_trash", kerosene = "h42_kero", incense = "h42_incence",                                   generator = "h42_generator", smoky = "h42_smoked", crop = "h42_crop", stove_other = "h42_smoky_kitc") %>% 
  select(-c(concatenated, timepoint)) %>% 
  mutate(arm = case_when(arm == 'R' ~ 'Control', arm == 'Q' ~ 'Intervention', TRUE ~ arm))
```

Then, I will create four dataframes, one for the personal filters, other two for the duplicates and another one for the blank filters. I will also remove the unnecessary columns from these dataframes.

```{r}
#Creating the dataframe for personal sample filters, and deleting the other columns with unnecessary filter IDs
hapin_personal <- hapin %>% filter(filter_id %in% xrflist$filter_id) %>% 
  select(-c(blank_filterid, dup1_filterid, dup2_filterid))

#Creating the dataframe for duplicates in column 4, and deleting the other columns with unnecessary filter IDs, and renaming `dup1_filterid` to `filter_id`.
hapin_dup1 <- hapin %>% filter(dup1_filterid %in% xrflist$filter_id) %>% 
  select(-c(blank_filterid, filter_id, dup2_filterid)) %>% 
  rename(filter_id = "dup1_filterid")

#Creating the dataframe for duplicates in column 5, and deleting the other columns with unnecessary filter IDs, and renaming `dup2_filterid` to `filter_id`.
hapin_dup2 <- hapin %>% filter(dup2_filterid %in% xrflist$filter_id) %>% 
  select(-c(blank_filterid, filter_id, dup1_filterid)) %>% 
  rename(filter_id = "dup2_filterid")

#Creating the dataframe for blank filters, and deleting the other columns with unnecessary filter IDs, and renaming `blank_filterid` to `filter_id`.
hapin_blanks <- hapin %>% filter(blank_filterid %in% xrflist$filter_id) %>% 
  select(-c(filter_id, dup1_filterid, dup2_filterid)) %>% 
  rename(filter_id = "blank_filterid")
```

And now I will combine the personal and duplicates dataframes into a single one named `hapin_filters`.

```{r}
#Combining the `personal` and `dup` dataframes into a single one, using `bind_rows()`
hapin_filters <- bind_rows(hapin_personal, hapin_dup1, hapin_dup2)
```

Both the personal exposure sample and the blank dataframes are ready to be used for further analysis, so I saved them in RDS format.

```{r}
saveRDS(hapin_filters, file = here("data", "processed-data", "hapin_samples.rds"))
saveRDS(hapin_blanks, file = here("data", "processed-data", "hapin_blanks.rds"))
```

## Chemical species Concentrations data cleaning

The following sections will be long, but necessary to demonstrate how to clean all the data and make all the adjustments before getting the final data for analysis. First I will load the file with the chemical species concentrations, assigned to the `concentration` dataframe. I will also load the limit of detection values from all chemical species under the `lod` dataframe.

```{r}
#Loading the chemical species concentration data
concentration <- read_excel(here("data", "raw-data", "conc_unc_hapingt_emollinedo.xlsx"), sheet = "Concentration")

#Loading the chemical species limit of detection values
lod <- read_excel(here("data", "raw-data", "conc_unc_hapingt_emollinedo.xlsx"), sheet = "DL")

#having a glimpse of this dataframe
glimpse(concentration)
```

First things first, some of the observations contain brackets, which represent values below the detection limit and they have to be corrected in future steps. So first, here I will delete the brackets, then I will convert these variables to numeric type and round them to 4 decimal digits.

```{r}
#Remove brackets, convert to numeric, and round to 4 decimal places
concentration <- concentration %>%
  mutate(across(3:26, ~ round(as.numeric(str_remove_all(., "\\[|\\]")), 4))) %>% 
  rename(filter_id = "Filter")
```

Now, I will remove the filters that correspond to field blanks, since those will be used to adjust the sample filters. I will create a new data frame `blanks`, that contains those observations.

```{r}
#Create dataframe with the concentrations for the blank filters for each element
blanks <- concentration %>% filter(filter_id %in% hapin_blanks$filter_id) %>% 
  select(-c(filter_id, Type))

#Remove blank filters from the `concentration` dataframe
concentration <- concentration %>%
  anti_join(hapin_blanks, by = "filter_id")

#Calculate the median for each element, using the blank filter values
median_blanks <- blanks %>% select(Mg: Na) %>%
  map_dbl(median) %>% round(4)

median_blanks
```

Now, I will estimate which chemical species are representative, by filtering all the species with less than 50% of their values are below the detection limit.

```{r}
#Round the limit of detection values to 4 decimal places
lod <- lod %>% mutate(`Concentration (ug/cm2)` = round(`Concentration (ug/cm2)`, 4))
print(lod)

#Estimate the percentage of values below detection limit for each element
bdl <- concentration %>%
  summarise(
    percent_bdl_mg = sum(Mg < 0.0051) / n() * 100,
    percent_bdl_al = sum(Al < 0.0059) / n() * 100,
    percent_bdl_si = sum(Si < 0.0068) / n() * 100,
    percent_bdl_s = sum(S < 0.0026) / n() * 100,
    percent_bdl_k = sum(K < 0.0040) / n() * 100,
    percent_bdl_ca = sum(Ca < 0.0040) / n() * 100,
    percent_bdl_ti = sum(Ti < 0.0086) / n() * 100,
    percent_bdl_cr = sum(Cr < 0.0055) / n() * 100,
    percent_bdl_mn = sum(Mn < 0.0030) / n() * 100,
    percent_bdl_fe = sum(Fe < 0.0043) / n() * 100,
    percent_bdl_ni = sum(Ni < 0.0043) / n() * 100,
    percent_bdl_cu = sum(Cu < 0.0080) / n() * 100,
    percent_bdl_zn = sum(Zn < 0.0024) / n() * 100,
    percent_bdl_ga = sum(Ga < 0.0049) / n() * 100,
    percent_bdl_as = sum(As < 0.0062) / n() * 100,
    percent_bdl_se = sum(Se < 0.0181) / n() * 100,
    percent_bdl_cd = sum(Cd < 0.0932) / n() * 100,
    percent_bdl_in = sum(In < 0.1089) / n() * 100,
    percent_bdl_sn = sum(Sn < 0.1844) / n() * 100,
    percent_bdl_te = sum(Te < 0.2166) / n() * 100,
    percent_bdl_i = sum(I < 0.3149) / n() * 100,
    percent_bdl_pb = sum(Pb < 0.0324) / n() * 100,)

print(bdl)
```

Only 10 elements have less than 50% of their observations below the detection limit. A note: Na and Cl were seen in high and very variable concentrations, so those elements will also be removed from the analysis. Now, I will filter the remaining elements (variables).

```{r}
#Select only the filter ID, type and the 10 elements that have more than 50% of their data not below the detection limit
concentration <- concentration %>% select(c(filter_id, Type, Mg, Al, Si, S, K, Ca, Ti, Mn, Fe, Zn)) %>% 
  arrange(filter_id) #Also I decided to arrange them ascending based on the filter_id
```

Now, blank adjusting all the values, by substracting the median value for each element, obtained from the blank filters. Also, replace all values below the detection limit with (LOD/sqrt(2)).

```{r}
#Substract the median blank value for the elements that need blank adjustment
concentration <- concentration %>%
  mutate(Mg = Mg - 0.0112, S = S - 0.0044, K = K - 0.0014, Ca = Ca - 0.0048, Fe = Fe - 0.0038, Zn = Zn - 0.0045)

#Replace all values below the detection limit with DL/sqrt(2)
concentration <- concentration %>% 
    mutate(Mg = round(if_else(Mg < 0.0051, 0.0051 / sqrt(2), Mg), 4), Al = round(if_else(Al < 0.0059, 0.0059 / sqrt(2), Al), 4),
           Si = round(if_else(Si < 0.0068, 0.0068 / sqrt(2), Si), 4), S = round(if_else(S < 0.0026, 0.0026 / sqrt(2), S), 4),
           K = round(if_else(K < 0.0016, 0.0016 / sqrt(2), K), 4), Ca = round(if_else(Ca < 0.0040, 0.0040 / sqrt(2), Ca), 4),
           Ti = round(if_else(Ti < 0.0086, 0.0086 / sqrt(2), Ti), 4), Mn = round(if_else(Mn < 0.0030, 0.0030 / sqrt(2), Mn), 4),
           Fe = round(if_else(Fe < 0.0043, 0.0043 / sqrt(2), Fe), 4), Zn = round(if_else(Zn < 0.0024, 0.0024 / sqrt(2), Zn), 4))

#I will save these concentrations in a new dataframe `concentrations_ugcm2` that will be used for uncertainty estimations in future steps
concentrations_ugcm2 <- concentration %>% select(-Type)
```

And then, converting the concentrations from micrograms per square centimeter to micrograms per cubic meter using the equation: ug/m3 = (Concentration (ug/cm^2^) x Filter area (cm^2^))/Sample volume (m3). Where, the filter area is 0.7853 cm^2^ and the sample volume is 0.43 m^3^. Also, I need to input the BC concentrations from the `hapin_filters` dataframe.

```{r}
#Converting the concentrations from (ug/cm2) to (ug/m3)
concentration <- concentration %>%
  mutate(across(Mg:Zn, ~ round(.x * 0.7853 / 0.43, 4)))

#Now adding the BC concentrations value from the `hapin_filters` df
hapin_filters <- hapin_filters %>% arrange(filter_id) #First arrange in ascending order by filter_id
concentration <- bind_cols(concentration, hapin_filters$bc) %>% rename(BC = "...13") %>% select(-Type) #Join the BC concentrations to the `concentrations` df, also renaming the column to "BC" and deleting the 'Type' column
```

I will add the fueltype and arm variables to this dataframe, and then I'll save it in RDS format.

```{r}
#Add the `arm` and `fueltype` variables to the `concentration` dataframe
concentration <- concentration %>%
  mutate(arm = hapin_filters$arm, fueltype = hapin_filters$fueltype)

#Save this dataframe as RDS file
saveRDS(concentration, file = here("data", "processed-data", "concentration.rds"))
```

## Chemical species Uncertainties data processing

Now, it's time to make some adjustments to the uncertainties from all filters across all chemical species. First, loading the uncertainties data to the `uncertainty` data frame.

```{r}
uncertainty <- read_excel(here("data", "raw-data", "conc_unc_hapingt_emollinedo.xlsx"), sheet = "Uncertainty")
```

First, I will remove the chemical species with more than 50% of their values below detection limit, also remove the blank filter observations, sort by Filter ID and round the values to numbers with four decimal places.

```{r}
uncertainty <- uncertainty %>% select(c(Filter, Mg, Al, Si, S, K, Ca, Ti, Mn, Fe, Zn)) %>% #Select only the necessary variables
  rename(filter_id = "Filter") %>% #Rename the Filter column
  mutate_at(vars(Mg:Zn), ~ round(., 4)) %>% #Round the uncertainties to 4 decimal places
  arrange(filter_id) %>% #Arrange observations based on Filter ID
  anti_join(hapin_blanks, by = "filter_id") #Remove blank filters

#Have a glimpse of the data
glimpse(uncertainty)
```

Now, the uncertainties from the concentration values that were LOD adjusted have to be reestimated to (5/6)\*LOD. To do this I created a list that inputs the LOD values for each element, and then using `mutate()`, I make the changes only to the values that follow the specified condition.

```{r}
#Create a list for the detection limits for each element
dl <- c(Mg = 0.0051, Al = 0.0059, Si = 0.0068, S = 0.0026, K = 0.0016, 
        Ca = 0.0040, Ti = 0.0086, Mn = 0.0030, Fe = 0.0043, Zn = 0.0024)

# Recalculate the values in df based on the conditions
uncertainty <- uncertainty %>%
  mutate(across(all_of(names(dl)),
                ~ round(if_else(concentrations_ugcm2[[cur_column()]] < dl[[cur_column()]], (5/6) * dl[[cur_column()]], .), 4),
                .names = "{.col}"))
```

Finally, for all the chemical species, the uncertainties have to be reestimated so they correctly reflect the uncertainty for each variable used in the calculations. In this case, input the XRF-estimated concentration, filter area and filter volume uncertainties using the law of propagation of uncertainties. Also, the purpose of this section is to convert the uncertainties from ug/cm2 to ug/m3 to correctly reflect the units of the concentrations.

```{r}
# Define the variables to be recalculated
variables <- c('Mg', 'Al', 'Si', 'S', 'K', 'Ca', 'Ti', 'Mn', 'Fe', 'Zn')

# Recalculate the values using the given equation
uncertainty <- uncertainty %>%
  mutate(across(all_of(variables), ~ round( #Here, selecting all the variables to be re-calculated
    sqrt((0.7853/0.43)^2 * .^2 + (concentrations_ugcm2[[cur_column()]]/0.43)^2 * 0.05^2 + (-concentrations_ugcm2[[cur_column()]]*0.7853/0.43^2)^2 * (0.05 * 0.43)^2), 4), .names = "{.col}")) #Input the equation
```

Since the BC concentrations will also be included, I need to estimate the BC uncertainties. These estimations require an additional file that will be assigned to the `atn` dataframe.

```{r}
#Load file to estimate BC uncertainties
atn <- read_csv(here("data", "raw-data", "ATN_summary_SA.csv"))

#Explore the structure of the file
glimpse(atn)
```

To estimate the BC concentrations, I used an instrument called Sootscan, that calculates transmissometry as attenuation before (pre) and after (post) sampling the filters. These values summarize the pre- and post- ATN values for all the filters, but I have to estimate the total attenuation for all sample filters with the equation: LogN(post-ATN/pre-ATN)\*100.

```{r}
#Estimate total attenuation by sample
atn <- atn %>% mutate(ATN = round(log(post_ATN/pre_ATN) *100, 2))
```

And finally, estimate the BC uncertainties using an equation also derived from the lag of propagation of uncertainties.

```{r}
#Estimate the BC uncertainties
uncertainty <- uncertainty %>% 
  mutate(BC = round((1370 * sqrt((atn$ATN)^2 * 0.000005^2 + 0.000079^2 * 1.16^2)), 4))
```

I will also add the arm and fueltype variables to this dataframe, and I'll save it as RDS format.

```{r}
#Add the `arm` and `fueltype` variables to the `uncertainty` dataframe
uncertainty <- uncertainty %>%
  mutate(arm = hapin_filters$arm, fueltype = hapin_filters$fueltype)

#Save this dataframe as RDS file
saveRDS(uncertainty, file = here("data", "processed-data", "uncertainty.rds"))
```