### 2024-03-15 products/manuscript/supplement/Supplementary.qmd

---
title: "Supplementary Material"
subtitle: "Project: Part 3"
author: Erick E. Mollinedo
date: '`r format(Sys.Date(), "%B %d, %Y")`'
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
---

```{r include=FALSE}
library(tidyverse)
library(tidymodels)
library(here)
library(readr)
```

# Supplementary Material

The following document includes some of the results from the exploratory data analysis.

## Supplementary Results

@fig-conc shows the summary of concentrations for each chemical species categorized by study arm. It is observed a similar pattern observed from the categorization by type of fuel, where Mg, Mn, Ca, K, BC, Ti, Si and S have statistically significant differences in concentrations (shown as ****).

![Chemical species concentrations by study arm](../../../results/figures/boxplot-arm.jpg){#fig-conc}

In terms of the categorical exposures @fig-cat summarizes the responses from all the exposures. It is observed that for the incence and generator variables there were just a few or none observations that recorded being exposed to these sources. Meanwhile, stove usage was the exposure most recorded.

![Summary of categorical exposures](../../../results/figures/cat-exposures.jpg){#fig-cat}

The following figures show the distribution of the concentrations of Carbon monoxide (@fig-co) and PM~2.5~ (@fig-pm), which were considered as covariates (CO) or independent variables (PM~2.5~) to be included in the modelling process with the categorical exposures.

![Carbon Monoxide concentrations by type of fuel](../../../results/figures/co.jpg){#fig-co}

![PM2.5 concentrations by type of fuel](../../../results/figures/pm25.jpg){#fig-pm}

The following tables show the results when conducting the generalized linear models using a single or multiple variables for the chemical species that can be interpreted by sources of exposure.

```{r echo=FALSE}
#Open the hapin_merged df

hapin_merged <- read_csv(here("data", "processed-data", "hapin-merged.csv"))

# Create a generalized linear regression model specification using gamma distribution (This will be used for all the models)
model_spec <- linear_reg(mode = "regression") %>%
  set_engine("glm", family = Gamma(link = "log"))

###########To predict the BC variable##############
# Create the recipe for preprocessing
recipeBC_red <- recipe(BC ~ kerosene + stove_other + fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowBC_red <- workflow() %>%
  add_recipe(recipeBC_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitBC_red <- workflowBC_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsBC_red <- fitBC_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsBC_red)

###########To predict the Ca variable##############
# Create the recipe for preprocessing
recipeCa_red <- recipe(Ca ~ fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowCa_red <- workflow() %>%
  add_recipe(recipeCa_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitCa_red <- workflowCa_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsCa_red <- fitCa_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsCa_red)

###########To predict the K variable##############
# Create the recipe for preprocessing
recipeK_red <- recipe(K ~ smoke + stove_other + fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowK_red <- workflow() %>%
  add_recipe(recipeK_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitK_red <- workflowK_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsK_red <- fitK_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsK_red)

###########To predict the Mg variable##############
# Create the recipe for preprocessing
recipeMg_red <- recipe(Mg ~ kerosene + fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowMg_red <- workflow() %>%
  add_recipe(recipeMg_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitMg_red <- workflowMg_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsMg_red <- fitMg_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsMg_red)

###########To predict the Mn variable##############
# Create the recipe for preprocessing
recipeMn_red <- recipe(Mn ~ fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowMn_red <- workflow() %>%
  add_recipe(recipeMn_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitMn_red <- workflowMn_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsMn_red <- fitMn_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsMn_red)

###########To predict the S variable##############
# Create the recipe for preprocessing
recipeS_red <- recipe(S ~ fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowS_red <- workflow() %>%
  add_recipe(recipeS_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitS_red <- workflowS_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsS_red <- fitS_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsS_red)

###########To predict the Fe variable##############
# Create the recipe for preprocessing
recipeFe_red <- recipe(Fe ~ stove, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowFe_red <- workflow() %>%
  add_recipe(recipeFe_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitFe_red <- workflowFe_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsFe_red <- fitFe_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsFe_red)

###########To predict the Si variable##############
# Create the recipe for preprocessing
recipeSi_red <- recipe(Si ~ fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowSi_red <- workflow() %>%
  add_recipe(recipeSi_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitSi_red <- workflowSi_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsSi_red <- fitSi_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsSi_red)

###########To predict the Ti variable##############
# Create the recipe for preprocessing
recipeTi_red <- recipe(Ti ~ fueltype, data = hapin_merged) %>%
  step_dummy(all_nominal(), -all_outcomes())

# Create a workflow
workflowTi_red <- workflow() %>%
  add_recipe(recipeTi_red) %>%
  add_model(model_spec)

# Fit the model to the data
fitTi_red <- workflowTi_red %>%
  fit(data = hapin_merged)

# Extract and print the results of the model
resultsTi_red <- fitTi_red %>%
  pull_workflow_fit() %>%
  tidy()

print(resultsTi_red)
```