### 2024-04-19 products/manuscript/supplement/Supplementary.qmd

---
title: "Supplementary Material"
author: Erick E. Mollinedo
date: '`r format(Sys.Date(), "%B %d, %Y")`'
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
---

```{r include=FALSE}
library(tidyverse)
library(tidymodels)
library(here)
library(readr)
library(gt)
```

# Supplementary Material

The following document includes some of the results from the exploratory data analysis.

## Supplementary Results

@fig-conc shows the summary of concentrations for each chemical species categorized by study arm. It is observed a similar pattern observed from the categorization by type of fuel, where Mg, Mn, Ca, K, BC, Ti, Si and S have statistically significant differences in concentrations (shown as \*\*\*\*).

![Chemical species concentrations by study arm](../../../results/figures/boxplot-arm.png){#fig-conc}

In terms of the categorical exposures @fig-cat summarizes the responses from all the exposures. It is observed that for the incence and generator variables there were just a few or none observations that recorded being exposed to these sources. Meanwhile, stove usage was the exposure most recorded.

![Summary of categorical exposures](../../../results/figures/cat-exposures.jpg){#fig-cat}

The following figures show the distribution of the concentrations of Carbon monoxide (@fig-co) and PM~2.5~ (@fig-pm), which were considered as covariates (CO) or independent variables (PM~2.5~) to be included in the modelling process with the categorical exposures.

![Carbon Monoxide concentrations by type of fuel](../../../results/figures/co.png){#fig-co}

![PM2.5 concentrations by type of fuel](../../../results/figures/pm25.png){#fig-pm}

The following tables show the model summaries for the final models selected. In this case the linear regression model for the PM~2.5~ mass concentrations and the gamma GLM model for the BC mass concentrations.

```{r echo=FALSE}
#Load the `hapin_merged` dataframe
hapin_merged <- read_rds(here("data", "processed-data", "hapin-final.rds"))

# Split the data into training and testing sets
rngseed = 1234
set.seed(rngseed) # for reproducibility
data_split <- initial_split(hapin_merged, prop = 0.8)
train_data <- training(data_split)
test_data <- testing(data_split)

#Create formulas
pm25formula <- pm25 ~ Mg + Al + Si + S + K + Ca + Ti + Mn + Fe + Zn + BC
bcformula <- BC ~ smoke + coil + trash + kerosene + smoky + crop + stove_other + fueltype

#Set model specifications
lin_mod <- linear_reg() %>% set_engine("lm") #Linear
glm_mod <- linear_reg(mode = "regression") %>% #GLM gamma
  set_engine("glm", family = Gamma(link = "log"))

## ----Final model PM2.5: Linear regression---- ##
linpm_wflow <- workflow() %>% #Workflow
	add_model(lin_mod) %>% 
	add_formula(pm25formula)

linpm_fit <- linpm_wflow %>% fit(data = train_data) #Fit the model

## ----Final model BC: GLM Gamma---- ##
glmbc_wflow <- workflow() %>% #Workflow
  add_model(glm_mod) %>% 
  add_formula(bcformula)

glmbc_fit <- glmbc_wflow %>% fit(data = train_data) #Fit the model

# Extract and print the results of the models
results_pmlin <- linpm_fit %>% #PM2.5 Linear
  extract_fit_parsnip() %>%
  tidy()

results_bcglm <- glmbc_fit %>% #BC GLM
  extract_fit_parsnip() %>%
  tidy()

#Print the results
print(results_pmlin)
print(results_bcglm)
```

The following figures illustrate the observed vs predicted values plots for both PM~2.5~ and BC models, and the residual plots.

![Observed vs Predicted values of the PM2.5 models](../../../results/figures/obs-pred_pm25.png)

![PM2.5 Models residuals](../../../results/figures/residuals-pm25.png)

![Observed vs Predicted values of the BC models](../../../results/figures/obs-pred_bc.png)

![BC models residuals](../../../results/figures/residuals-bc.png)