### 2024-04-25 project-review-Hu.qmd

---
title: Project Review for Mollinedo
author: Xueyan Hu
date: 04-24-2024
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Elemental characterization of personal filter samples from the Household Air Pollution Intervention Network (HAPIN) trial for Source Apportionment

Name of project author(s): Erick E. Mollinedo

Name of project reviewer: Xueyan Hu

# Instructions

Write your comments and feedback below for each section/component of the project. The goal should be to help the author improve their project. Make comments as constructive and actionable as possible. You can provide both criticism and praise.

For each component, pick one summary statement by deleting the ones that do not apply and keeping only the one that you think most closely summarizes a given component.

Make sure your final document compiles/knits into a readable, well-formatted html document.

Delete any sections/text of this template that are not part of your final review document. (Including these instructions.)

# Specific project content evaluation
Evaluate the different parts of the project by filling in the sections below.

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

There is an detailed background description and it is easy to understand the current condition of household air pollution. And the transition from introduction to the hypothesis/question is natural and they are highly related.

### Summary assessment (PICK ONE, DELETE THE OTHERS)

* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The questions raised are clear enough and they provide me a rough frame that how the dataset would be organized, what variables it might include, and what kind of analysis the author would apply to the dataset.

### Summary assessment

* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

The source is provided and the calculating equation is also listed. Although I could have a rough understanding for the raw data, it will be a little better if a codebook is created (I am sorry if I miss this file.)

### Summary assessment

* source and overall structure of data somewhat explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

The data cleaning process is well organized, all of the cleaning steps are well illustrated, and the processed data looks good to me.

### Summary assessment

* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

The analysis methods are appropriate for the data in this project according to the type of outcome. And there is a specific model related to environmental analysis involved in this project, different from what models we usually learn during the course. I think it is suitable for data analysis of this project.

### Summary assessment

* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

Tables are used for presenting the result. A summary is written firstly and then it is divided into several parts for further description. Figure 1 is impressive to me because tt allows me to intuitively find the relationship between variables, especially in the condition that many variables are included it seems really important. The metircs are proper and supplement file is well used.

### Summary assessment

* results are very well presented

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

The study findings are properly discussed with strengths and limitations mentioned, such as linear model can fit the data well and low diversity in answers for categorical variables could lead to bias. And I think both are reasonable.

### Summary assessment

* strong, complete and clear discussion

## Further comments

_Overall, this is a mature, highly completed and meaningful project. From data collection to data cleaning, to the use and supplement of all interpreted figures in the model, every part shows that the project is well organized, and we can also see how much thinking and time the author spent on the project.But there is only one small thing that I would like to mention. There is a folder called R which is created in the template by Dr. Handel. The author haven't used any file in this folder, so I think it could be deleted._

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

The project is well structured and files are well labeled in the folders. So it is quite easy for me to find all of the relative materials. All of the files have reasonable name. But there is only one folder named 'R' left which should be deleted. And under the result file, maybe 'large file' folder and 'output' folder could also be deleted.

### Summary assessment

* mostly clear, but some confusing parts (e.g. useless files, things in the wrong folders)

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

Yes, I think reading through data analysis qmd file, coupled with the contents of the manuscript, I can understand the whole process and don't have much difficulty. But the process of data cleaning is kind of complicated, and it needs to be reviewed back and forth for better understanding, but this is only my own situation.

### Summary assessment

* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

I have tried all the codes involved in this project and they are fully reproducible. But I haven't found the instruction for reproducing this whole project in the supplement file.
I think it is "fully reproducible without issues" but it will be better if some instructions could be provided.

### Summary assessment

* fully reproducible without issues

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

The overall study is thorough, and both linear and logistic models are used to fit the dataset and the performances are compared, evaluated and discussed in different files. And the result mentioning predictors and the better model that can answer the questions raised above.

### Summary assessment

* strong level of thoroughness

## Further comments

_I just have a small question. Because only simple linear model and glm are fit in this project, plus the number of variables involved in the data is relatively larger, I wonder if more models could be tried to fit the data to compare the performance and maybe the overfitting problem could also be considered in the discussion part._