### 2024-04-26 project-review-Kaggwa.qmd_Erick.qmd
Review by Patrick
Including review by Patrick K.

---
title: Project Review Template 
author: Patrick Kaggwa
date: 04-24-2024
format: 
  html:
    toc: true
    toc-depth: 3
    number-sections: true
---

# Overview

Title of project: Elemental characterization of personal filter samples from the Household Air Pollution Intervention Network (HAPIN) trial for Source Apportionment

Name of project author(s): Erick E. Mollinedo

Name of project reviewer: Patrick Kaggwa

# Instructions

Write your comments and feedback below for each section/component of the project. The goal should be to help the author improve their project. Make comments as constructive and actionable as possible. You can provide both criticism and praise.

For each component, pick one summary statement by deleting the ones that do not apply and keeping only the one that you think most closely summarizes a given component.

Make sure your final document compiles/renders into a readable, well-formatted html document.

Delete any sections/text of this template that are not part of your final review document. (Including these instructions.)

# Specific project content evaluation.

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

* Context of the Project: 
The project provides a solid background on household air pollution (HAP) in low- and middle-income countries (LMICs). While the background offers a decent summary of previous work, it could be strengthened with more explicit referencing. The project is reasonably well-placed within the existing scientific and public health discourse.
The introduction effectively outlines why the project was undertaken by highlighting the significance of HAP, the limited research on SA in LMICs.

* Areas of improvement 
*Conciseness: 
Some parts of the introduction and background sections contain detailed explanations that could be condensed for brevity without sacrificing clarity. 
* Citations:
While citations are provided throughout the text, there are instances where additional citations could strengthen the claims made. 
* Integration of Pilot Study Findings: 
The introduction briefly mentions the pilot study conducted in Guatemala and Rwanda as part of the HAPIN trial, but it could be more explicit in integrating the findings of this pilot study into the rationale for the current research.

### Summary assessment (PICK ONE, DELETE THE OTHERS)

* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments
The questions and hypotheses that the project aims to address are fairly clear and well-described.

### Summary assessment

* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments
Data Source is well described and the text provides a detailed description of how the data was collected and processed.

#Codebook provided: 
The code book is loaded well

### Summary assessment
* source and overall structure of data well explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

The exploratory data analysis appears sound and is thoroughly documented. The goals for each step were evident, and the coding and procedures were straightforward and easily understandable.

### Summary assessment

* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments
The analysis appears to be done properly, with appropriate methods used for each component of the analysis. The code is well-explained and provides clear justification for the chosen approaches.
### Summary assessment

* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

Overall, the presentation of the results is well-executed, striking a good balance between textual explanation and visual representation, which enhances the clarity and accessibility of the findings.

### Summary assessment

* results are very well presented

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

Overall, the discussion effectively addresses the study findings, acknowledges both strengths and limitations, and provides a proper interpretation of the results, contributing to a comprehensive understanding of the research outcomes.

### Summary assessment

* strong, complete and clear discussion

## Further comments

Overall, the project demonstrates a strong understanding of the subject matter and presents the findings in a coherent and thoughtful manner.

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments
The project is well structured but it still has default files and folders which should removed.

### Summary assessment
* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

Overall, while the project is reasonably well-documented and understandable.

### Summary assessment

* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

Everything run smoothly and it reproducible.

### Summary assessment

* fully reproducible without issues

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

Overall the study is thorough.

### Summary assessment

* strong level of thorougness

## Further comments

_I feel justice was done but remove all unused folders and files as they still exist_